---
title: "MidPoint Deliverable"
author: "Winnie Chen"
date: "5/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r echo = F, include=FALSE}
library(knitr)
library(dplyr)
library(readxl)

source("scripts/analysis.R")

source("scripts/summary_info.R")

source("scripts/agg_table.R")

source("scripts/charts_wen.R")

source("scripts/analysis_graphnicole.R")
```

## Intro
Our group project primarily focuses on unemployment because right now, during this pandemic, unemployment has been at an all time high; however, we want to analyze what other factors could affect unemployment. We analyzed and see how unemployment relates to education, basic demographics, and poverty. 

There were several datasets that was used in this project, such as unemployment by state and area, as well as education levels from primary school to Master's degree. Our project mainly focuses on three dataset: poverty, basic demographics, and education. The poverty dataset compares how poverty is spread throughout the U.S., allowing us to examine the relationship between unemployment and poverty and see what states and counties have the highest prevalence. As for the education and basic demographics dataset, it contains the unemployment rate based on the education qualification of adults and race.

```{r echo=FALSE, include=FALSE}
education_df <- read_xls("data/education_usda.xls", range = "Education 1970 to 2018!A5:AU3288")
colnames(education_df) <- gsub(' ', '_', colnames(education_df))
colnames(education_df) <- gsub(',', '', colnames(education_df))
colnames(education_df) <- gsub("'", '', colnames(education_df))
colnames(education_df) <- gsub('-', '_', colnames(education_df))

poverty_df <- read_xls("data/PovertyEstimates_usda.xls", range = "Poverty Data 2018!A5:AB3198")
colnames(poverty_df) <- gsub(' ', '_', colnames(poverty_df))
colnames(poverty_df) <- gsub(',', '', colnames(poverty_df))

unemployment_df <- read_xls("data/Unemployment_usda.xls", range = "Unemployment Med HH Income!A8:CJ3283")
colnames(unemployment_df) <- gsub(' ', '_', colnames(unemployment_df))
colnames(unemployment_df) <- gsub(',', '', colnames(unemployment_df))

education_pov_unemployment_df <- education_df %>%
  left_join(poverty_df, by = c("FIPS_Code" = "FIPStxt")) %>%
  left_join(unemployment_df, by = c("FIPS_Code" = "FIPStxt"))

county_summary <- get_summary_info(education_pov_unemployment_df)
```

## Summary Information
In 2018, the mean percent of adults living in a given county in the U.S.
without a highschool diploma was 
`r round(county_summary$mean_county_no_diploma_rate, digits = 2)`%,
while the mean percent of adults with a bachelor's degree or more was
`r round(county_summary$mean_county_bach_rate, digits = 2)`%. The average
unemployment rate and poverty rate for U.S. counties were
`r round(county_summary$mean_county_unemployment, digits = 2)`% and
`r round(county_summary$mean_pov_rate, digits = 2)`% respectively.
The county with the lowest poverty rate had
`r round(county_summary$min_pov_rate, digits = 2)`% of its population
living in poverty, while the county with the highest poverty rate had
`r round(county_summary$max_pov_rate, digits = 2)`% of its population living
in poverty.

## Average Unemployment Rate With Different Levels of Education
```{r eval=T, echo=F}
unemployment_kag_df <- read.csv("data/unemployment_data_us_kaggle.csv",
                                stringsAsFactors = F)

aggregate_table <- get_agg_table(unemployment_kag_df)
kable(aggregate_table, digits = 2)
```

needs to be added in

## Chart 1: Unemployment vs Poverty

The purpose of this graph is to show the correlation between the unemployment of each state and the number of people in poverty. Each dot represents a state and their unemployment rate and how that affects the people in poverty. A scatter plot would be the best plot for this type of data since it makes it easy for the user to see how it correlates with each other. 

```{r echo=FALSE}
unemployment_vs_poverty <- 
  read.csv("data/Unemployement_Poverty_2016_kaggle.csv",
                                    stringsAsFactors = FALSE)

unemployment_by_state <- unemployment_vs_poverty %>%
  group_by(State) %>%
  summarise(
    unemployment_rate = sum(Unemployment_rate_2016, na.rm = TRUE),
    people_in_poverty = sum(POVALL_2016, na.rm = TRUE)) %>%
  arrange(desc(unemployment_rate))

visualization(unemployment_by_state, "State")

```

This chart reveals that there's a positive correlation between unemployment and people in poverty. The lower the unemployment, the lower the number of people in poverty. It also shows that many states don't have high unemployment and people in poverty, however, there are no states that has exactly zero. In addition, we see a few outliers in the chart, two specifically that are way out of line with the rest of the states. Regardless of the two outliers, we can see that there is a positive correlation between unemployment and people in poverty.

## Chart 2: Unemployment vs Education Level

```{r echo=FALSE}
unemployment_df <- read.csv("data/unemployment_data_us_kaggle.csv", stringsAsFactors = FALSE)

df_1 <- unemployment_df %>%
  group_by(Year) %>%
  summarize(avg_unemployment_rate = mean(Primary_School, na.rm = T)) %>%
  mutate(education_level = "Primary School")

df_2 <- unemployment_df %>%
  group_by(Year) %>%
  summarize(avg_unemployment_rate = mean(High_School, na.rm = T)) %>%
  mutate(education_level = "High School")

df_3 <- unemployment_df %>%
  group_by(Year) %>%
  summarize(avg_unemployment_rate = mean(Associates_Degree, na.rm = T)) %>%
  mutate(education_level = "Associates Degree")

df_4 <- unemployment_df %>%
  group_by(Year) %>%
  summarize(avg_unemployment_rate = mean(Professional_Degree, na.rm = T)) %>%
  mutate(education_level = "Professional Degree")

summary <- rbind(df_1, df_2, df_3, df_4)

new_df <- summary %>%
  filter(Year != "2020")

stacked_bar_chart(new_df)
```

I included this stacked bar chart because it compares the unemployment rates of each education level by year. This is easier to see the differences. The chart reveals the average unemployment rate for each education level by year. One thing I observed from this was that the discrepancy between a primary school education level and a professional school education level got smaller as the years passed. In addition, the chart descended as the years passed, which means that unemployment levels decreased.

## Chart 3: Unemployment Demographics

```{r echo=FALSE}
df <- read.csv("data/unemployment_data_us_kaggle.csv", stringsAsFactors = FALSE)

df <- mutate(df, avg_unemployment = (White + Black + Asian + Hispanic) / 4)

plot_filter2 <- df %>%
  filter(Year != "2020") %>%
  group_by(Month) %>%
  arrange(Year) %>%
  select(Year, Month, White, Black, Asian, Hispanic, avg_unemployment)

graph_function(plot_filter2)
```

This chart reveals that among black, white, hispanic, and asian people, black people had a higher unemployment rate from 2010 - 2019. The green line shows the average unemployment rate and this allows us to see that asian and white people are below the average unemployment rate while hispanic people are close to average and black people are above average. This graph creates a visualization to see how unemployment rates were widespread in 2010 but have become closer to the same rate in 2019 although black people still have the highest rate of unemployment. 
