---
title: "Test"
author: "Winnie Chen"
date: "5/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r echo = F, include=FALSE}
library(knitr)

source("analysis.R")

source("scripts/summary_info.R")
```

## Intro
Fill in later

```{r echo=FALSE, include=FALSE}
education_df <- read.csv("data/education_usda.csv", stringsAsFactors = F)
colnames(education_df)[1] <- gsub('^...', '', colnames(education_df)[1])

poverty_df <- read.csv("data/PovertyEstimates_usda.csv")
colnames(poverty_df)[1] <- gsub('^...', '', colnames(poverty_df)[1])

unemployment_df <- read.csv("data/Unemployment_usda.csv")
colnames(unemployment_df)[1] <- gsub('^...', '', colnames(unemployment_df)[1])

education_poverty_unemployment_df <- education_df %>% 
  left_join(poverty_df, by = c("FIPS.Code" = "FIPStxt")) %>% 
  left_join(unemployment_df, by = c("FIPS.Code" = "FIPStxt"))

county_summary <- get_summary_info(education_poverty_unemployment_df)
```

## Summary Information
In 2018, the mean percent of adults living in a given county in the U.S.
without a highschool diploma was 
`r round(county_summary$mean_county_no_diploma_rate, digits = 2)`%,
while the mean percent of adults with a bachelor's degree or more was
`r round(county_summary$mean_county_bach_rate, digits = 2)`%. The average
unemployment rate and poverty rate for U.S. counties were
`r round(county_summary$mean_county_unemployment, digits = 2)`% and
`r round(county_summary$mean_pov_rate, digits = 2)`% respectively.
The county with the lowest poverty rate had
`r round(county_summary$min_pov_rate, digits = 2)`% of its population
living in poverty, while the county with the highest poverty rate had
`r round(county_summary$max_pov_rate, digits = 2)`% of its population living
in poverty.


